Quantile Regressions via Method of Moments with multiple fixed effects

### 2023-10-24 article/article.qmd
s
s

More recently, @mss2019 (MSS hereafter) proposed a methodology based on a conditional location-scale model similar to the one described in @he1997 and @zhao2020, for the estimation of quantile regressions models for panel data via a method of moments. This method allows individual fixed effects allowing to have heterogeneous effects on the entire conditional distribution of the outcome, rather constraining their effect to be a location shift only as in @canay2011, @koenker2004, and @lancaster2000.

In principle, under the assumption that data generating process behind the data is based on a multiplicative heteroskedastic process that is linear in parameters (@cameron2005; @mss2019; @he1997; @zhao2020), the effect of a variable $X$ on the $q_th$ quantile can be derived as the combination of a location effect, and scale effect moderated by the quantile of an underlying i.i.d. error. For statistical inference, MSS derives the asymptotic distribution of the estimator, suggesting the use of bootstrap standard errors, as well.

@@

The rest of the paper is restructured as follows. Section 2 presents the basic setup of the location-Scale model described in @he1997 and @zhao2020, tying the relationship between the standard quantile regression model, and the location and scale model. It also revisits MSS methodology, proposing alternative estimators for the standard errors based on the properties of GMM estimators and the empirical influence functions. It also shows that FWL theorem can be used to control for multiple fixed effects. Section 3 presents the results of a small simulation study and Section 4 illustrates the application of the proposed methods with two empirical examples. Seccion 5 concludes.

@@

Call $\theta=[ \beta' \ \ \gamma'  \ \ Q(\varepsilon|\tau)' ]'$ the set of coefficients that are identified in @eq-eq10, a just identified model.  And the functions $G_{k,l}$ and $S_{k.l}$ equal to:

$$\begin{aligned}
G_{k,l} &= N^{-1} \sum_{i=1}^N \frac{\partial h_{k,i}}{\partial \theta_l} \Big|_{\hat\theta} \ \forall k,l \in [1,2,3] \\
S_{k,l} &= N^{-1} \sum_{i=1}^N h_{k,i} h_{l,i}' \Big|_{\hat\theta} \ \forall k,l \in [1,2,3]
\end{aligned}
$$

where $\theta_1 = \beta$, $\theta_2 = \gamma$ and $\theta_3 = Q(\varepsilon,\tau)$.

Then 
$$
\sqrt{N}(\hat\theta - \theta) \xrightarrow{d} N(0, G^{-1} S G^{-1})
$$

with, $G$ and $S$ defined as:

$$\begin{aligned}
G &= \begin{bmatrix}
G_{1,1} &       0 & 0       \\
G_{2,1} & G_{2,2} & 0       \\
G_{3,1} & G_{3,2} & G_{3,3} 
\end{bmatrix}  \text{ and }
S &= \begin{bmatrix}
S_{1,1} & S_{1,2} & S_{1,3}    \\
S_{2,1} & S_{2,2} & S_{2,3}  \\
S_{3,1} & S_{3,2} & S_{3,3} 
\end{bmatrix}
\end{aligned}
$$

Because

$$
Var(\hat\theta) = \frac{1}{N} \left(
    \frac{1}{N} 
    \sum_{i=1}^N  \lambda_i(\theta) \lambda_i(\theta)'  
\right) = \frac{1}{N^2} 
\lambda(\theta)'\lambda(\theta)
$$

$$
\lambda_i(\theta) = \bar G(\theta)^{-1} h_i (x_i,\theta)
$$

Finally, the variance covariance of $\hat \beta(\tau)$ will be given by:

$$Var(\hat\beta(\tau)) = \Xi Var(\hat\theta) \Xi'$$

@@

$$
\begin{aligned}
    \Xi &= I(k), \hat q(\tau) I(k), \hat \gamma \\
    \Xi &= \left[ \begin{array}{cccc|cccc|}
           1 & 0 & \dots & 0 &\hat q_(\tau) & 0 & ... & 0  & \hat\gamma_0 \\
           0 & 1 & \dots & 0 &0& \hat q_(\tau) & ... & 0 & \hat\gamma_1 \\
          \vdots & \vdots & \dots & 0 &...&...& ... & 0             & ...\\
           0 & 0 & \dots & 1 &0 & 0 & ... & \hat q(\tau)   & \hat\gamma_k
        \end{array} \right]         
\end{aligned}
$$

>>>>

More recently, @mss2019 (MSS hereafter) proposed a methodology based on a conditional location-scale model similar to the one described in @he1997 and @zhao2000, for the estimation of quantile regressions models for panel data via a method of moments. This method allows individual fixed effects allowing to have heterogeneous effects on the entire conditional distribution of the outcome, rather constraining their effect to be a location shift only as in @canay2011, @koenker2004, and @lancaster2000.

In principle, under the assumption that data generating process behind the data is based on a multiplicative heteroskedastic process that is linear in parameters (@cameron2005, @mss2019, @he1997 and @zhao2000 ), the effect of a variable $X$ on the $q_th$ quantile can be derived as the combination of a location effect, and scale effect moderated by the quantile of an underlying i.i.d. error. For statistical inference, MSS derives the asymptotic distribution of the estimator, suggesting the use of bootstrap standard errors, as well.

@@

The rest of the paper is restructured as follows. Section 2 presents the basic setup of the location-Scale model described in @he1997 and @zhao2000, tying the relationship between the standard quantile regression model, and the location and scale model. It also revisits MSS methodology, proposing alternative estimators for the standard errors based on the properties of GMM estimators and the empirical influence functions. It also shows that FWL theorem can be used to control for multiple fixed effects. Section 3 presents the results of a small simulation study and Section 4 illustrates the application of the proposed methods with two empirical examples. Seccion 5 concludes.

@@

Call $\theta=[ \beta' \ \ \gamma'  \ \ Q(\varepsilon|\tau)' ]'$ the set of coefficients that are identified in @eq-eq10, a just identified model.  And the function $h_i$ is a vector function that stacks all the moments at the individual level described in @eq-eq10. Then $\hat\theta$ follows a normal distribution with mean $\theta$ and variance-covariance matrix $V(\theta)$ that is estimated as:
$$
\hat{V}(\hat\theta)=\frac{1}{N} 
\bar G(\hat\theta)^{-1} 
\left( \frac{1}{N} \sum_{i=1}^N h_i h_i'  \Big|_{\theta=\hat\theta} \right) 
\bar G(\hat\theta)^{-1} 
$$

where the inner product is the moment covariance matrix, and $\bar{G}(\theta)$ is the Jacobian matrix of the moment equations (@eq-eq10) evaluated at $\hat\theta$.

$$\bar{G}(\theta) =-\frac{1}{N} \sum_{i=1} \frac{\partial h_i}{\partial \theta'} \Big|_{\theta=\hat\theta}$$

This is equivalent to the Eicker-White Heteroskedastic-Consistent estimator for least-squares estimators.

In this framework, the quantile regression coefficients, a combination of the location-scale-quantile estimates, will follow a normal distribution with mean $\beta(\tau) = \beta+Q(\varepsilon,\tau)\gamma$ and variance-covariance matrix equal to:

$$\hat{V}(\beta(\tau)) = \Xi \hat{V}(\hat\theta) \Xi'
$$

@@

$$\Xi = [ I(k), \hat Q(\varepsilon|\tau) \cdot I(k), \hat \gamma ]
$$

with $I(k)$ being an identity matrix of dimension $k$ (number of explanatory variables in $X$ including the constant).

The different types of Standard errors estimation, thus, depend on the assumptions imposed for the estimation of $V(\theta)$.

### 2023-10-24 article/article.tex
s
s

\title{Quantile Regressions via Method of Moments with multiple fixed
effects}
\author{Fernando Rios-Avila \and Leonardo Siles \and Gustavo
Canavire-Bacarreza}
\date{2023-10-23}

\begin{document}
\maketitle
\begin{abstract}
This paper proposes a new method to estimate quantile regressions with
multiple fixed effects. The method expands on the strategy proposed by
Machado and Santos Silva (2019), allowing for multiple fixed effects,
and providing various alternatives for the estimation of Standard
errors. We provide Monte Carlo simulations to show the finite sample
properties of the proposed method in the presence of two sets of fixed
effects. Finally, we apply the proposed method to estimate
\textbf{something interesting}
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

More recently, Machado and Santos Silva (2019) (MSS hereafter) proposed
a methodology based on a conditional location-scale model similar to the
one described in He (1997) and (\textbf{zhao2020?}), for the estimation
of quantile regressions models for panel data via a method of moments.
This method allows individual fixed effects allowing to have
heterogeneous effects on the entire conditional distribution of the
outcome, rather constraining their effect to be a location shift only as
in Canay (2011), Koenker (2004), and Lancaster (2000).

In principle, under the assumption that data generating process behind
the data is based on a multiplicative heteroskedastic process that is
linear in parameters (Cameron and Trivedi (2005); Machado and Santos
Silva (2019); He (1997); (\textbf{zhao2020?})), the effect of a variable
\(X\) on the \(q_th\) quantile can be derived as the combination of a
location effect, and scale effect moderated by the quantile of an
underlying i.i.d. error. For statistical inference, MSS derives the
asymptotic distribution of the estimator, suggesting the use of
bootstrap standard errors, as well.

@@

The rest of the paper is restructured as follows. Section 2 presents the
basic setup of the location-Scale model described in He (1997) and
(\textbf{zhao2020?}), tying the relationship between the standard
quantile regression model, and the location and scale model. It also
revisits MSS methodology, proposing alternative estimators for the
standard errors based on the properties of GMM estimators and the
empirical influence functions. It also shows that FWL theorem can be
used to control for multiple fixed effects. Section 3 presents the
results of a small simulation study and Section 4 illustrates the
application of the proposed methods with two empirical examples. Seccion
5 concludes.

@@

Call \(\theta=[ \beta' \ \ \gamma' \ \ Q(\varepsilon|\tau)' ]'\) the set
of coefficients that are identified in Equation~\ref{eq-eq10}, a just
identified model. And the functions \(G_{k,l}\) and \(S_{k.l}\) equal
to:

\[\begin{aligned}
G_{k,l} &= N^{-1} \sum_{i=1}^N \frac{\partial h_{k,i}}{\partial \theta_l} \Big|_{\hat\theta} \ \forall k,l \in [1,2,3] \\
S_{k,l} &= N^{-1} \sum_{i=1}^N h_{k,i} h_{l,i}' \Big|_{\hat\theta} \ \forall k,l \in [1,2,3]
\end{aligned}
\]

where \(\theta_1 = \beta\), \(\theta_2 = \gamma\) and
\(\theta_3 = Q(\varepsilon,\tau)\).

Then \[
\sqrt{N}(\hat\theta - \theta) \xrightarrow{d} N(0, G^{-1} S G^{-1})
\]

with, \(G\) and \(S\) defined as:

\[\begin{aligned}
G &= \begin{bmatrix}
G_{1,1} &       0 & 0       \\
G_{2,1} & G_{2,2} & 0       \\
G_{3,1} & G_{3,2} & G_{3,3} 
\end{bmatrix}  \text{ and }
S &= \begin{bmatrix}
S_{1,1} & S_{1,2} & S_{1,3}    \\
S_{2,1} & S_{2,2} & S_{2,3}  \\
S_{3,1} & S_{3,2} & S_{3,3} 
\end{bmatrix}
\end{aligned}
\]

Because

\[
Var(\hat\theta) = \frac{1}{N} \left(
    \frac{1}{N} 
    \sum_{i=1}^N  \lambda_i(\theta) \lambda_i(\theta)'  
\right) = \frac{1}{N^2} 
\lambda(\theta)'\lambda(\theta)
\]

\[
\lambda_i(\theta) = \bar G(\theta)^{-1} h_i (x_i,\theta)
\]

Finally, the variance covariance of \(\hat \beta(\tau)\) will be given
by:

\[Var(\hat\beta(\tau)) = \Xi Var(\hat\theta) \Xi'\]

@@

\[
\begin{aligned}
    \Xi &= I(k), \hat q(\tau) I(k), \hat \gamma \\
    \Xi &= \left[ \begin{array}{cccc|cccc|}
           1 & 0 & \dots & 0 &\hat q_(\tau) & 0 & ... & 0  & \hat\gamma_0 \\
           0 & 1 & \dots & 0 &0& \hat q_(\tau) & ... & 0 & \hat\gamma_1 \\
          \vdots & \vdots & \dots & 0 &...&...& ... & 0             & ...\\
           0 & 0 & \dots & 1 &0 & 0 & ... & \hat q(\tau)   & \hat\gamma_k
        \end{array} \right]         
\end{aligned}
\]

>>>>

\title{Quantile Regressions via Method of Moments with multiple fixed
effects}
\author{Fernando Rios-Avila \and Leonardo Siles \and Gustavo
Canavire-Bacarreza}
\date{2023-10-24}

\begin{document}
\maketitle
\begin{abstract}
This paper proposes a new method to estimate quantile regressions with
multiple fixed effects. The method expands on the strategy proposed by
Machado and Santos Silva (2019), allowing for multiple fixed effects,
and providing various alternatives for the estimation of Standard
errors. We provide Monte Carlo simulations to show the finite sample
properties of the proposed method in the presence of two sets of fixed
effects. Finally, we apply the proposed method to estimate
\textbf{something interesting}
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi

@@

More recently, Machado and Santos Silva (2019) (MSS hereafter) proposed
a methodology based on a conditional location-scale model similar to the
one described in He (1997) and Zhao (2000), for the estimation of
quantile regressions models for panel data via a method of moments. This
method allows individual fixed effects allowing to have heterogeneous
effects on the entire conditional distribution of the outcome, rather
constraining their effect to be a location shift only as in Canay
(2011), Koenker (2004), and Lancaster (2000).

In principle, under the assumption that data generating process behind
the data is based on a multiplicative heteroskedastic process that is
linear in parameters (Cameron and Trivedi (2005), Machado and Santos
Silva (2019), He (1997) and Zhao (2000) ), the effect of a variable
\(X\) on the \(q_th\) quantile can be derived as the combination of a
location effect, and scale effect moderated by the quantile of an
underlying i.i.d. error. For statistical inference, MSS derives the
asymptotic distribution of the estimator, suggesting the use of
bootstrap standard errors, as well.

@@

The rest of the paper is restructured as follows. Section 2 presents the
basic setup of the location-Scale model described in He (1997) and Zhao
(2000), tying the relationship between the standard quantile regression
model, and the location and scale model. It also revisits MSS
methodology, proposing alternative estimators for the standard errors
based on the properties of GMM estimators and the empirical influence
functions. It also shows that FWL theorem can be used to control for
multiple fixed effects. Section 3 presents the results of a small
simulation study and Section 4 illustrates the application of the
proposed methods with two empirical examples. Seccion 5 concludes.

@@

Call \(\theta=[ \beta' \ \ \gamma' \ \ Q(\varepsilon|\tau)' ]'\) the set
of coefficients that are identified in Equation~\ref{eq-eq10}, a just
identified model. And the function \(h_i\) is a vector function that
stacks all the moments at the individual level described in
Equation~\ref{eq-eq10}. Then \(\hat\theta\) follows a normal
distribution with mean \(\theta\) and variance-covariance matrix
\(V(\theta)\) that is estimated as: \[
\hat{V}(\hat\theta)=\frac{1}{N} 
\bar G(\hat\theta)^{-1} 
\left( \frac{1}{N} \sum_{i=1}^N h_i h_i'  \Big|_{\theta=\hat\theta} \right) 
\bar G(\hat\theta)^{-1} 
\]

where the inner product is the moment covariance matrix, and
\(\bar{G}(\theta)\) is the Jacobian matrix of the moment equations
(Equation~\ref{eq-eq10}) evaluated at \(\hat\theta\).

\[\bar{G}(\theta) =-\frac{1}{N} \sum_{i=1} \frac{\partial h_i}{\partial \theta'} \Big|_{\theta=\hat\theta}\]

This is equivalent to the Eicker-White Heteroskedastic-Consistent
estimator for least-squares estimators.

In this framework, the quantile regression coefficients, a combination
of the location-scale-quantile estimates, will follow a normal
distribution with mean \(\beta(\tau) = \beta+Q(\varepsilon,\tau)\gamma\)
and variance-covariance matrix equal to:

\[\hat{V}(\beta(\tau)) = \Xi \hat{V}(\hat\theta) \Xi'
\]

@@

\[\Xi = [ I(k), \hat Q(\varepsilon|\tau) \cdot I(k), \hat \gamma ]
\]

with \(I(k)\) being an identity matrix of dimension \(k\) (number of
explanatory variables in \(X\) including the constant).

The different types of Standard errors estimation, thus, depend on the
assumptions imposed for the estimation of \(V(\theta)\).
